---
title: "Average mpg of mtcars Dataset"
author: "Salomon Villatoro"
date: "6/1/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(tinytex)
library(tidytext)
library(knitr)

```

This is the Rmd file with the requested datable that shows the average mile per 
gallon of each car make in the mtcars dataset.

The code I used below highlights how I thought about answering this problem.

First, I used regular expressions to extract all car makes from the car models
in the dataset.

```{r mtcars}
mtcars_avg <- mtcars %>% 
  rownames_to_column(var = "car") %>% 
  mutate(make = str_extract(car,"[A-Z][a-z]+"))
```

Next, I grouped my previously saved table by make, and calculated the mean mpg
for each make using the tapply function in the apply family.
```{r cars}
cars <- data.frame(tapply(mtcars_avg$mpg, mtcars_avg$make,mean)) %>% 
  rownames_to_column(var = "Make") %>% 
  rename(mpg = "tapply.mtcars_avg.mpg..mtcars_avg.make..mean.") %>% 
  mutate(mpg = round(mpg, 2))
```
Finally, I used knitr's kable function to clean up the table for presentation.
```{r kable}

kable(list(cars[1:11,], cars[12:22,]),
      caption = "Average miles per gallon for each car make",
      booktabs = TRUE)

```
